Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-04-06 aggexplain/abstract.tex
typo fixing

>>>>

\begin{contributions}{What is the impact of aggregating instance-level explanations?}
\item Histograms can be used to aggregate instance-level explanations and compare subsets effectively.
\item Aggregating instance-level explanations significantly outperforms inspecting individual explanations or unassisted aggregation for detecting biases in the data.
\item Inspecting individual instance-level explanations can be misleading and hurts detecting biases in the data.
\item Aggregated instance-level explanations allow to detect biases equally well as in a table without explanations while being more scalable.
\end{contributions}

### 2018-04-06 aggexplain/experiment.tex
typo fixing

@@

\subsection{Study Design}
We study four conditions which result from combining different representations with the inclusion or exclusion of explanations.
The different representations are:

>>>>

\begin{figure}
\centering
\includegraphics[width=\linewidth]{aggexplain/full_table_expl}
\caption[The full interface illustrating the table view.]{
The full interface illustrating the table view showing individual instances.
The user is comparing the model's prediction of ``high'' house prices (orange) to the prediction of ``low'' prices (purple).
The feature ``Living Area'' is ordered by ascending values and the user hovers over the cell with the value ``605''.
}
\label{figs:full_view}
\end{figure}

@@

\vspace*{-0.5em}
\subsection{Study Design}
\vspace*{-0.5em}
We study four conditions which result from combining different representations with the inclusion or exclusion of explanations.
The different representations are:

### 2018-04-06 aggexplain/relatedwork.tex
much more text
typo fixing

\subsection{Effectiveness of instance-level explanations}
When introducing their algorithm, LIME (Ribeiro~\etal~\cite{DBLP:journals/corr/RibeiroSG16, anchors:aaai18,2016arXiv161105817T}), the authors conducted experiments to show the effectiveness of their method.
However, instance-level explanations were only inspected individually and not in aggregate form.

@@

Stumpf~\etal~\cite{harmful} found that under some circumstances, explanations can be harmful to the end user, by invoking false confidence.
This is on one hand due to the user extrapolating from few instance-level explanations, making their mental model seem correct.
And, on the other hand, trust in the machine learning model \emph{overrides} their initial intuition: ``I guess this thing knows more than me. The system knows more than me. I'll accept [the diagnosis]''.
We could confirm both of those findings in our experiments.

\subsection{Detecting biases using visual analytics methods}
Hohman~\etal~\cite{2018arXiv180106889H} identifies detecting biased data as one of their five use cases for visual analytics for machine learning.
However, their examples focus on work that only looks at the data without the help of machine learning models~\cite{facets} or simple models where humans adjust the thresholds of the model manually~\cite{wattenberg}.

>>>>

\subsection{Effectiveness of Instance-level Explanations}
When introducing their algorithm, LIME (Ribeiro~\etal~\cite{DBLP:journals/corr/RibeiroSG16, anchors:aaai18,2016arXiv161105817T}), the authors conducted experiments to show the effectiveness of their method.
However, instance-level explanations were only inspected individually and not in aggregate form.

@@

Stumpf~\etal~\cite{harmful} found that under some circumstances, explanations can be harmful to the end user, by invoking false confidence.
This is on one hand due to the user extrapolating from few instance-level explanations, making their mental model seem correct.
And, on the other hand, trust in the machine learning model \emph{overrides} their initial intuition: ``I guess this thing knows more than me. The system knows more than me. I'll accept [the diagnosis]''.
The study investigated inspecting individual instances one after the other, however,
in our experiments we could confirm both of those findings even when showing instances in tables.

\subsection{Detecting Biases Using Visual Analytics Methods}
Hohman~\etal~\cite{2018arXiv180106889H} identifies detecting biased data as one of their five use cases for visual analytics for machine learning.
However, their examples focus on work that only looks at the data without the help of machine learning models~\cite{facets} or simple models where humans adjust the thresholds of the model manually~\cite{wattenberg}.

### 2018-04-06 aggexplain/results.tex
typo fixing

\begin{figure}
\centering
% \begin{subfigure}{0.45\linewidth}
\includegraphics[width=0.45\linewidth]{aggexplain/stats/tvss}%
% \caption{Trust and sense in the biased model.}
% \label{subfig:tvss}
% \end{subfigure}%
% \hfill%
% \begin{subfigure}{0.45\linewidth}
% \includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-all}%
% \caption{Change between both models.}
% \label{subfig:tvss-chg}
% \end{subfigure}%
\caption[Trust and whether the model makes sense.]{
A kernel density estimation of the responses of participants on a five-point Likert scale
about how much they trust the biased model and whether it makes sense to them.
Note, that the majority of participants did not detect the bias of the model.
% (b) shows the change of those values when comparing the unbiased model to biased model.
% A positive value indicates that the response was higher in the unbiased model.
}
\label{figs:tvss}
\end{figure}

\subsection{Bias Detection and Trust}

When comparing how much participants trust a particular model and whether they think that model makes sense, one can see that those responses are typically correlated (see for example Figure~\ref{figs:tvss} with a Pearson correlation coefficient of $0.759$ and Spearman rank-order correlation coefficient of $0.745$).
This also extends to plain-text answers, which allow to detect whether participants correctly found the bias in the correct data set unambiguously.
Participants were very verbose about their findings, if they found something:
``\emph{It has higher accuracy so should be more trustworthy than the other one. However some of the results don't make sense to me. Maybe this is just an atypical property market.}'';
``\emph{It is accurate, yet the predictions do not make much sense. Higher quality houses having a larger amount of low priced houses, percentage-wise? More rooms, area, or stories resulting in lower prices? The logic does not work out.}'';
``\emph{larger houses are valued lower than others which are smaller}''.

\begin{figure}
\centering
\begin{subfigure}{0.45\linewidth}
\includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-correct}%
\caption{\centering Bias~detected~and correct~preference.}
\label{subfig:tvss-chg-a}
\end{subfigure}%
\hfill%
\begin{subfigure}{0.45\linewidth}
\includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-change}%
\caption{\centering Bias~detected~but no~preference~change.}
\label{subfig:tvss-chg-b}
\end{subfigure}%
\caption[Change in trust and whether the model makes sense.]{
How participants changed their responses comparing the unbiased model to the biased model.
A positive value indicates that the response was higher in the unbiased model.
(a) shows the case when participants detected the bias and subsequently preferred the unbiased model.
(b) shows the case when participants detected the bias but still chose the biased model.
}
\label{figs:tvss-chg}
\end{figure}

However, the above mentioned correlation is not perfect.
This is likely due to some participants not being convinced, that their correct discovery of the flaw in the data is enough that the corresponding model cannot be trusted:
``\emph{If the data says it's true, then it's true I suppose and it's more trustworthy than my common sense.}'';
``\emph{I feel like the results of [the biased model] where strange even though they where correct according to the dataset.}'' (sic);
``\emph{I'm drawn to trusting the model which was more accurate even though it didn't entirely make sense to me.}''.

@@

In order to investigate this hypothesis further, we can look at the number of interactions of the participants performing the tasks.
We can see in Figure~\ref{figs:table_interaction} that participants engaged with the table view significantly more when no explanations were present.
This might be an example of Hullman~\etal~\cite{6064986}, which states that information visualization might benefit from visual difficulties, as is the case with a table without any further help from the interface about what to look at, since people are forced to interact more with the visualization.

>>>>

\begin{figure}
\centering
% \begin{subfigure}{0.45\linewidth}
\includegraphics[width=0.45\linewidth]{aggexplain/stats/tvss}%
% \caption{Trust and sense in the biased model.}
% \label{subfig:tvss}
% \end{subfigure}%
% \hfill%
% \begin{subfigure}{0.45\linewidth}
% \includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-all}%
% \caption{Change between both models.}
% \label{subfig:tvss-chg}
% \end{subfigure}%
\caption[Trust and whether the model makes sense.]{
A kernel density estimation of the responses of participants on a five-point Likert scale
about how much they trust the biased model and whether it makes sense to them.
Note, that the majority of participants did not detect the bias of the model.
% (b) shows the change of those values when comparing the unbiased model to biased model.
% A positive value indicates that the response was higher in the unbiased model.
}
\label{figs:tvss}
\vspace*{-0.5em}
\end{figure}

\subsection{Bias Detection and Trust}
\vspace*{-0.5em}
When comparing how much participants trust a particular model and whether they think this model makes sense, one can see that those responses are typically correlated (see for example Figure~\ref{figs:tvss} with a Pearson correlation coefficient of $0.759$ and Spearman rank-order correlation coefficient of $0.745$).
This also extends to plain-text answers, which allow to detect whether participants correctly found the bias in the correct data set unambiguously.
Participants were very verbose about their findings, if they found something:
``\emph{It has higher accuracy so should be more trustworthy than the other one. However some of the results don't make sense to me. Maybe this is just an atypical property market.}'';
``\emph{It is accurate, yet the predictions do not make much sense. Higher quality houses having a larger amount of low priced houses, percentage-wise? More rooms, area, or stories resulting in lower prices? The logic does not work out.}'';
``\emph{larger houses are valued lower than others which are smaller}'' (sic).

\begin{figure}[t]
\centering
\begin{subfigure}{0.49\linewidth}
\includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-correct}%
\caption{\centering Bias~detected~and correct~preference.}
\label{subfig:tvss-chg-a}
\end{subfigure}%
\hfill%
\begin{subfigure}{0.49\linewidth}
\includegraphics[width=\linewidth]{aggexplain/stats/tvss-chg-change}%
\caption{\centering Bias~detected~but no~preference~change.}
\label{subfig:tvss-chg-b}
\end{subfigure}%
\caption[Change in trust and whether the model makes sense.]{
How participants changed their responses comparing the unbiased model to the biased model.
A positive value indicates that the response was higher in the unbiased model.
(a) shows the case when participants detected the bias and subsequently preferred the unbiased model.
(b) shows the case when participants detected the bias but still chose the biased model.
}
\label{figs:tvss-chg}
\end{figure}

However, the above mentioned correlation is not perfect.
This is likely due to some participants not being convinced, that their correct discovery of the flaw in the data is enough that the corresponding model cannot be trusted:
``\emph{If the data says it's true, then it's true I suppose and it's more trustworthy than my common sense.}'';
``\emph{I feel like the results of [the biased model] where strange even though they where correct according to the dataset.}'';
``\emph{I'm drawn to trusting the model which was more accurate even though it didn't entirely make sense to me.}'' (sic).

@@

In order to investigate this hypothesis further, we can look at the number of interactions of the participants performing the tasks.
We can see in Figure~\ref{figs:table_interaction} that participants engaged with the table view significantly more when no explanations were present.
This might be an example of Hullman~\etal~\cite{6064986}, who state that information visualization might benefit from visual difficulties, since people are forced to interact more with the visualization. This seems to be the case for a table, without any further help from the interface about what to look at.

### 2018-04-06 explainer/motivation.tex
typo fixing

We use the term \textit{model diagnostics} to indicate the steps necessary for a domain expert or a model developer to semantically validate the decisions made by a model using their domain knowledge. In this section we outline the different goals for a user when using a model diagnostic interface and provide an overview of the implementation of the resulting workflow~(Figure~\ref{figs:workflow}). 
The workflow was derived through a long term collaboration among visual analytic researchers and model developers and domain experts in the medical field, specifically in the application scenario of hospital visits. The over-arching goal in this scenario is to use predictive modeling for reducing patient wait time and optimizing the hospital resources needed for admitted patients.

>>>>

We use the term \textit{Model Diagnostics} to indicate the steps necessary for a domain expert or a model developer to semantically validate the decisions made by a model using their domain knowledge. In this section we outline the different goals for a user when using a model diagnostic interface and provide an overview of the implementation of the resulting workflow~(Figure~\ref{figs:workflow}). 
The workflow was derived through a long term collaboration among visual analytic researchers and model developers and domain experts in the medical field, specifically in the application scenario of hospital visits. The over-arching goal in this scenario is to use predictive modeling for reducing patient wait time and optimizing the hospital resources needed for admitted patients.

### 2018-04-06 explainer/relwork.tex
typo fixing

\subsection{Human-in-the-Loop Inspection of Classifiers}%Classification Models}

Researchers have recently demonstrated how human interventions can help in greater accuracy in construction of classifiers, when compared with a purely automated approach~\cite{tam2017analysis}. In this work, Tam \etal used information theory to show how soft knowledge of model developers can be encoded in decision trees, and they advocate a tighter integration between human and machine-centric processes for model development. The goals for integrating visual analytic techniques and classification methods fall broadly into three categories, as proposed by Liu \etal~\cite{liu2017towards}: i) model understanding, ii) model diagnosis, and iii) model refinement. Our proposed diagnostic workflow~(Figure~\ref{figs:workflow}) encompasses the goals of understanding model behavior and diagnosing the model decision space for enabling data scientists and domain experts to generate insights about potential inadequacies in the data and in the model quality. The refinement step is an obvious action as a result of these insights, but is outside the scope of our work.

>>>>

\vspace*{-1em}
\subsection{Human-in-the-Loop Inspection of Classifiers}%Classification Models}
\vspace*{-0.5em}
Researchers have recently demonstrated how human interventions can help in greater accuracy in construction of classifiers, when compared with a purely automated approach~\cite{tam2017analysis}. In this work, Tam \etal used information theory to show how soft knowledge of model developers can be encoded in decision trees, and they advocate a tighter integration between human and machine-centric processes for model development. The goals for integrating visual analytic techniques and classification methods fall broadly into three categories, as proposed by Liu \etal~\cite{liu2017towards}: i) model understanding, ii) model diagnosis, and iii) model refinement. Our proposed diagnostic workflow~(Figure~\ref{figs:workflow}) encompasses the goals of understanding model behavior and diagnosing the model decision space for enabling data scientists and domain experts to generate insights about potential inadequacies in the data and in the model quality. The refinement step is an obvious action as a result of these insights, but is outside the scope of our work.

### 2018-04-06 explainer/ui.tex
typo fixing

The fourth column shows the odds ratio of the subset on a logarithmic scale (Figure~\ref{figs:expl_main}H).
Whiskers indicate its confidence interval.
Odds ratio is a popular metric for determining effectiveness in evidence based medicine and clinical trials.
It is computed by comparing the subset explained by the given explanation with the full set of current data items.
This way we can detect whether an explanation describes a consistent subset of instances or if the subset appears like a random sample.
With this the odds ratio is:
\[
\frac{p_e / n_e}{p_t / n_t}
\]
where $p_e$ and $n_e$ is the ratio of positive and negative items respectively in the explanation subset and $p_t$ and $n_t$ is the ratio of those items in the remaining data set.
The confidence interval of the odds ratio is then computed as:
\[
\exp{\left(\log{(\text{odds ratio})} \; \pm \; 1.96 \sqrt{P_e^{-1} + N_e^{-1} + P_t^{-1} + N_t^{-1}}\right)}
\]
where $P$ and $N$ are the actual number of positive and negative items in the explanation subset $e$ and the remaining data $t$.

>>>>

The fourth column shows the odds ratio of the subset on a logarithmic scale (Figure~\ref{figs:expl_main}H).
Whiskers indicate its confidence interval.
Odds ratio is a popular metric for determining effectiveness in evidence based medicine and clinical trials.
It is computed by comparing the subset explained by the given explanation with the full set of current data items.
This way we can detect whether an explanation describes a consistent subset of instances or if the subset appears like a random sample.
With this the odds ratio is:
\vspace*{-0.5em}
\[
\frac{p_e / n_e}{p_t / n_t}
\]
\vspace*{-0.5em}
where $p_e$ and $n_e$ is the ratio of positive and negative items respectively in the explanation subset and $p_t$ and $n_t$ is the ratio of those items in the remaining data set.
The confidence interval of the odds ratio is then computed as:
\vspace*{-0.5em}
\[
\exp{\left(\log{(\text{odds ratio})} \; \pm \; 1.96 \sqrt{P_e^{-1} + N_e^{-1} + P_t^{-1} + N_t^{-1}}\right)}
\]
\vspace*{-0.5em}
where $P$ and $N$ are the actual number of positive and negative items in the explanation subset $e$ and the remaining data $t$.

### 2018-04-06 figs/workflow_intro.tex
much more text
typo fixing

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{figs/workflow_large}
\vspace*{-1.75em}
\caption{
The proposed \textit{model diagnostic} workflow extends the conventional \textit{model building} workflow in machine learning for enabling domain experts to reason about the semantic validity of the decisions made by any model through multiple linked visualizations.
This ultimately helps to improve data acquisition and model generation processes belonging to the original workflow.
}
\vspace*{-0.75em}
\label{figs:workflow_intro}
\end{figure*}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{figs/workflow_large}
\caption[The model diagnostic workflow.]{
The proposed \textit{Model Diagnostic} workflow extends the conventional \textit{Model Building} workflow in machine learning for enabling domain experts to reason about the semantic validity of the decisions made by any model through multiple linked visualizations.
This ultimately helps to improve data acquisition and model generation processes belonging to the original workflow.
}
\label{figs:workflow_intro}
\end{figure}

### 2018-04-06 infuse/introduction.tex
typo fixing

% !TEX root = ../featureselection.tex
The visualization research community has usually focused on developing techniques and systems to support the analysis of datasets, with limited analysis of the relationship between datasets and the construction of models on top of them. However, there are a growing number of data scientists interested in more than just interpreting their data: they want to understand their data and predictive probabilities associated with them. Providing visual support for this kind of task has become important as many existing applications on the market and in scientific settings need to solve problems that are predictive in nature, e.g. prediction of customer behavior, diseases, drug effectiveness.

Predictive modeling is defined as the process of developing a mathematical tool or model that generates an accurate prediction \cite{kuhn2013applied}. However, building an accurate predictive model is far from trivial. First, modelers must construct cohorts, or distinct groups, to divide their datasets into cases and controls. Then, they must use a feature construction technique to define the feature vector. Next, they must define the parameters for cross-validation to ensure the results are statistically valid and robust. Then, they need to choose a feature selection algorithm to extract the informative features and include them in a model. And finally, they need to choose a classifier to evaluate the predictiveness of the model. For each of these decisions, there are a variety of techniques for cohort construction, feature construction, cross-validation, features selection, and classification to choose from, and there are currently no systematic guidelines to decide which algorithms are most appropriate for which types of datasets. Making the wrong choices can cause predictive models to fail. Kuhn and John argue that many predictive models fail because, ``predictive modelers often only explore relatively few models when searching for predictive relationships [...] due to either modeler's preference for, or knowledge of, or expertise in, only a few models or the lack of available software that would enable them to explore a wide range of techniques" \cite{kuhn2013applied}. We use these current limitations as motivation to research how visual analytics may improve the process of predictive modeling.

>>>>

% !TEX root = ../featureselection.tex
The visualization research community has usually focused on developing techniques and systems to support the analysis of data sets, with limited analysis of the relationship between data sets and the construction of models on top of them. However, there are a growing number of data scientists interested in more than just interpreting their data: they want to understand their data and predictive probabilities associated with them. Providing visual support for this kind of task has become important as many existing applications on the market and in scientific settings need to solve problems that are predictive in nature, \eg, prediction of customer behavior, diseases, drug effectiveness.

Predictive modeling is defined as the process of developing a mathematical tool or model that generates an accurate prediction \cite{kuhn2013applied}. However, building an accurate predictive model is far from trivial. First, modelers must construct cohorts, or distinct groups, to divide their data sets into cases and controls. Then, they must use a feature construction technique to define the feature vector. Next, they must define the parameters for cross-validation to ensure the results are statistically valid and robust. Then, they need to choose a feature selection algorithm to extract the informative features and include them in a model. And finally, they need to choose a classifier to evaluate the predictiveness of the model. For each of these decisions, there are a variety of techniques for cohort construction, feature construction, cross-validation, features selection, and classification to choose from, and there are currently no systematic guidelines to decide which algorithms are most appropriate for which types of data sets. Making the wrong choices can cause predictive models to fail. Kuhn and John argue that many predictive models fail because, ``predictive modelers often only explore relatively few models when searching for predictive relationships [...] due to either modeler's preference for, or knowledge of, or expertise in, only a few models or the lack of available software that would enable them to explore a wide range of techniques" \cite{kuhn2013applied}. We use these current limitations as motivation to research how visual analytics may improve the process of predictive modeling.

### 2018-04-06 infuse/motivation.tex
typo fixing

\subsection{Predictive Modeling in Healthcare}
\label{sec:motivation_healthcare}
Predictive modeling is a common and important methodology used
in medical informatics and healthcare research.
For instance, it can be used to detect diseases in patients early
before they progress \cite{bellazzi2008predictive} and to
personalize treatment guidelines to understand which populations
will benefit from an intervention \cite{jensen2012mining}.
In order to derive such insights and build successful predictive models,
it is common for healthcare researchers to implement, evaluate,
and compare many models with different parameters and algorithms.
A common workflow for predictive models is a 5-step process,
illustrated in Figure~\ref{fig:pipeline}:
(1) cohort construction, (2) feature construction, (3) cross-validation,
(4) feature selection, and (5) classification.
There are currently few tools that support this complex
workflow for predictive modelers.

@@

\subsection{Running Example: Diabetes Prediction}
\label{sec:running_example}
In order to make our contributions concrete, we
utilize a running example from our case study.
Our case study involves a team of four clinical researchers interested
in using predictive modeling on a longitudinal database of
electronic medical records. The research team consisted of one MD researcher with a background in emergency medicine, and three PhD researchers with backgrounds in healthcare analytics.
Their database features over 300,000 patients from a major
healthcare provider in the United States.
The team is interested in building a predictive model to predict if a patient
is at risk of developing diabetes, a chronic disease of high blood sugar levels
that causes serious health complications.

>>>>

\subsection{Predictive Modeling in Health Care}
\label{sec:motivation_healthcare}
Predictive modeling is a common and important methodology used
in medical informatics and health care research.
For instance, it can be used to detect diseases in patients early
before they progress \cite{bellazzi2008predictive} and to
personalize treatment guidelines to understand which populations
will benefit from an intervention \cite{jensen2012mining}.
In order to derive such insights and build successful predictive models,
it is common for health care researchers to implement, evaluate,
and compare many models with different parameters and algorithms.
A common workflow for predictive models is a 5-step process,
illustrated in Figure~\ref{fig:pipeline}:
(1) cohort construction, (2) feature construction, (3) cross-validation,
(4) feature selection, and (5) classification.
There are currently few tools that support this complex
workflow for predictive modelers.

@@

\subsection{Running Example: Diabetes Prediction}
\label{sec:running_example}
In order to make our contributions concrete, we
utilize a running example from our case study.
Our case study involves a team of four clinical researchers interested
in using predictive modeling on a longitudinal database of
electronic medical records. The research team consisted of one MD researcher with a background in emergency medicine, and three PhD researchers with backgrounds in health care analytics.
Their database features over 300,000 patients from a major
health care provider in the United States.
The team is interested in building a predictive model to predict if a patient
is at risk of developing diabetes, a chronic disease of high blood sugar levels
that causes serious health complications.

### 2018-04-06 infuse/related_work.tex
typo fixing

Several approaches to feature selection and dimensionality reduction, in general, exist in visualization.
The early work of Guo~\cite{Guo2003} introduced the idea of
visualizing relationships between features sets.
His system is based on an interactive matrix view where rows and
columns represent features and the cells are colored according to
feature similarity (calculated as entropy and ${\chi}^2$).
The matrix is automatically sorted to allow selection of subspaces
(feature subsets) where data shows interesting clusters.
Visual hierarchical dimension reduction \cite{wang2003interactive}
allows detection and grouping of similar features as well.
The technique is based on a hierarchical clustering algorithm
which clusters dimensions in terms of their similarity and present
them in a \textit{sunburst} visualization \cite{yang2003interactive}.
Users can interactively choose an aggregation level and use the aggregated
dimensions to display data with the reduced set of dimensions.
Johansson and Johansson \cite{Johansson2009} present an integrated environment based on
\textit{parallel coordinates} visualization where the number and order of
dimensions (axes) presented at any time is guided by a ranking algorithm
that takes into account associations as well as intrinsic interestingness of
each feature to interactively choose how many features to visualize.
Similar in spirit is the \textit{rank-by-feature} framework \cite{seo2005rank} in which
the data features are organized, ranked and visualized in
1D and 2D visual representations
(e.g. histograms, bar charts and scatterplots).
The user can for instance inspect a matrix of feature pairs,
ranked by one of the available ranking functions, and single
out those that show interesting associations.
A similar mechanism is also used in \textit{scagnostics} \cite{wilkinson2005graph} a
quality metric approach \cite{bertini2011quality} that ranks axis pairs according to the pattern/shape they create in a scatterplot visualization.

@@

Visualization has also been used to aid in the creation of predictive models, not only in the selection of features that might be helpful in constructing such models. Visual construction and assessment of decision tree models have been the subject of a good number of works in the field. Ankerst \emph{et al.}, introduced the idea of using pixel-based visualization as a way to manually construct decision trees by giving the user the ability to observe class distributions within each node and to interactively select splitting points \cite{ankerst1999visual, ankerst2000towards}. A similar idea is proposed in \textit{PaintingClass} a visualization technique to manually build a decision tree through interaction of parallel coordinates and multidimensional scaling techniques to identify coherent groups of multidimensional data \cite{Teoh:2003:PIC:956750.956837}. More recently, \textit{BaobabView} has been presented as a system to inspect and validate a classification model through a tree representation. The paper presents a thorough analysis of the number of tasks that visualization can support in this area and how they are covered by the proposed system \cite{van2011baobabview}.

While all the aforementioned systems focus largely on decision trees, visualization has been used in other classification and regression systems that leverage other prediction models. The \textit{iVisClassifier} \cite{choo2010ivisclassifier} for instance uses \textit{linear discriminant analysis (LDA)}, a supervised dimensionality reduction method, to project multidimensional data in a scatterplot visualization taking into account information provided by the data labels. The technique allows to visually link the high-dimensional structure to the low-dimensional representation and build clusters. The clusters are then used to classify new data that is progressively introduced into the system to refine the model. Steed \emph{et al.}, in their \textit{cyclone trend analysis} provide a parallel coordinates visualization that leverage computational analysis to identify features with high predictive power in stepwise regression tasks and allows to build predictive models for multidimensional climate data \cite{steed2009guided, steed2009tropical}. Recently, a visual analytics system for regression analysis has been proposed by M\"uhlbacher and Piringer \cite{muhlbacher2013partition}. The system is more similar to our work in nature as it also focuses on the predictive power of feature sets and guides the user in the predictive modeling process. The main difference between this work and ours is our focus on classification rather than regression models and the use of multiple feature selection and classification models to better understand how features score across multiple models.

>>>>

Several approaches to feature selection and dimensionality reduction, in general, exist in visualization.
The early work of Guo~\cite{Guo2003} introduced the idea of
visualizing relationships between features sets.
His system is based on an interactive matrix view where rows and
columns represent features and the cells are colored according to
feature similarity (calculated as entropy and ${\chi}^2$).
The matrix is automatically sorted to allow selection of subspaces
(feature subsets) where data shows interesting clusters.
Visual hierarchical dimension reduction \cite{wang2003interactive}
allows detection and grouping of similar features as well.
The technique is based on a hierarchical clustering algorithm
which clusters dimensions in terms of their similarity and present
them in a \textit{sunburst} visualization \cite{yang2003interactive}.
Users can interactively choose an aggregation level and use the aggregated
dimensions to display data with the reduced set of dimensions.
Johansson and Johansson \cite{Johansson2009} present an integrated environment based on
\textit{parallel coordinates} visualization where the number and order of
dimensions (axes) presented at any time is guided by a ranking algorithm
that takes into account associations as well as intrinsic interestingness of
each feature to interactively choose how many features to visualize.
Similar in spirit is the \textit{rank-by-feature} framework \cite{seo2005rank} in which
the data features are organized, ranked and visualized in
1D and 2D visual representations
(e.g. histograms, bar charts and scatterplots).
The user can for instance inspect a matrix of feature pairs,
ranked by one of the available ranking functions, and single
out those that show interesting associations.
A similar mechanism is also used in \textit{scagnostics} \cite{wilkinson2005graph} a
quality metric approach \cite{bertini2011quality} that ranks axis pairs according to the pattern / shape they create in a scatterplot visualization.

@@

Visualization has also been used to aid in the creation of predictive models, not only in the selection of features that might be helpful in constructing such models. Visual construction and assessment of decision tree models have been the subject of a good number of works in the field. Ankerst~\etal, introduced the idea of using pixel-based visualization as a way to manually construct decision trees by giving the user the ability to observe class distributions within each node and to interactively select splitting points \cite{ankerst1999visual, ankerst2000towards}. A similar idea is proposed in \textit{PaintingClass} a visualization technique to manually build a decision tree through interaction of parallel coordinates and multidimensional scaling techniques to identify coherent groups of multidimensional data \cite{Teoh:2003:PIC:956750.956837}. More recently, \textit{BaobabView} has been presented as a system to inspect and validate a classification model through a tree representation. The paper presents a thorough analysis of the number of tasks that visualization can support in this area and how they are covered by the proposed system \cite{van2011baobabview}.

While all the aforementioned systems focus largely on decision trees, visualization has been used in other classification and regression systems that leverage other prediction models. The \textit{iVisClassifier} \cite{choo2010ivisclassifier} for instance uses \textit{linear discriminant analysis (LDA)}, a supervised dimensionality reduction method, to project multidimensional data in a scatterplot visualization taking into account information provided by the data labels. The technique allows to visually link the high-dimensional structure to the low-dimensional representation and build clusters. The clusters are then used to classify new data that is progressively introduced into the system to refine the model. Steed~\etal, in their \textit{cyclone trend analysis} provide a parallel coordinates visualization that leverage computational analysis to identify features with high predictive power in stepwise regression tasks and allows to build predictive models for multidimensional climate data \cite{steed2009guided, steed2009tropical}. Recently, a visual analytics system for regression analysis has been proposed by M\"uhlbacher and Piringer \cite{muhlbacher2013partition}. The system is more similar to our work in nature as it also focuses on the predictive power of feature sets and guides the user in the predictive modeling process. The main difference between this work and ours is our focus on classification rather than regression models and the use of multiple feature selection and classification models to better understand how features score across multiple models.

### 2018-04-06 infuse/system.tex
typo fixing

We provide a brief overview of data types utilized by the system. A predictive model, in our setting, is a model trained and validated with machine learning using a high number of features as an input to train the model. These features are the primary data items of \infuse. Each feature has a label representing the feature name (e.g. Diabetes), a category to which the feature belongs to (e.g. Diagnosis), and a subtype (e.g. Problem List, the health problems that led to the diagnosis).

>>>>

We provide a brief overview of data types utilized by the system. A predictive model, in our setting, is a model trained and validated with machine learning using a high number of features as an input to train the model. These features are the primary data items of \infuse. Each feature has a label representing the feature name (\eg, Diabetes), a category to which the feature belongs to (e.g. Diagnosis), and a subtype (\eg, Problem List, the health problems that led to the diagnosis).
